Filling in data gaps on refugee children: Statistical imputation for missing count data

### 2023-01-06 paper.qmd
caterpillar plots

### packages
library(tidyverse)
library(stringi)
library(gt)
library(ggplot2)
library(patchwork)
library(scales)

@@

>>>>

### packages
library(tidyverse)
library(stringi)
library(gt)
library(ggplot2)
library(patchwork)
library(scales)
library(ggdist)

@@

```{r}
##### check predictions #####

imputations <- imputations %>% 
  ungroup() %>%
  mutate(femaleProp = female/total,
         children = rowSums(select(., female_0_4, female_5_11, female_12_17, male_0_4, male_5_11, male_12_17))) %>% 
  mutate(childrenProp = children/total)

### by country of origin

## identify 3 large populations with missing age/sex data
t.dem.ageMissingTop9 <- dem_longMissing %>%
  ungroup() %>%
  filter(missing == "sexAge")%>%
  slice_max(total, n=9)

t.dem.ageMissingPlots <- dem_longMissing  %>% 
  filter(origin_iso3 == "VEN" & asylum_iso3 == "PER" & popType=="VDA" | origin_iso3 == "VNM" & asylum_iso3 == "CHN" |  origin_iso3 == "SYR" & asylum_iso3 == "SWE")

t.imp.ageMissingPlots <- imputations %>% 
  filter(origin_iso3 == "VEN" & asylum_iso3 == "PER" & popType=="VDA" | origin_iso3 == "VNM" & asylum_iso3 == "CHN" |  origin_iso3 == "SYR" & asylum_iso3 == "SWE")

# show for ifms
imputations.slides <- imputations %>% filter(imputation %in% c(1,2,3) & (origin_iso3 == "VEN" & asylum_iso3=="BOL" &popType == "VDA" | origin_iso3 == "SOM" & asylum_iso3=="UGA" | origin_iso3=="IDN" & asylum_iso3 == "PNG")) %>% arrange(origin_iso3, imputation) %>% select(imputation, origin_country, asylum_country, female_0_4:total)

#write.csv(imputations.slides, "ifms/imputations.slides.csv")

## Syria
# plot of uncertainty (full posterior distribution) 
# y: histogram
# x: proportion of female over imputations
# group: country of asylum

imputations.syr <- imputations %>% 
  filter(origin_iso3 == "SYR")

dem.syr.select <- dem_longMissing %>% filter(origin_iso3=="SYR", asylum_iso3 %in% c("SWE", "USA",  "MCO",  "LBN", "JOR", "GBR"))

p.imputations.syr.female <- ggplot(data = imputations.syr  ,
                            aes(y=reorder(asylum_country, femaleProp))) +
                    stat_histinterval(aes(x = femaleProp), slab_type  = "histogram")

p.imputations.syr.children <- ggplot(data = imputations.syr %>% filter(asylum_iso3 %in% c("SWE", "USA",  "MCO",  "LBN", "JOR", "GBR")),
      aes(y=reorder(asylum_country, childrenProp), x=childrenProp)) +
     stat_halfeye()

p.imputations.syr.reg.children <- ggplot(data = imputations.syr,
      aes(y=reorder(asylum_country, childrenProp), x=childrenProp)) +
     stat_gradientinterval() +
  facet_wrap(~ asylum_region, scales = "free")

p.imputations.syr.neighbour.children <- ggplot(data = imputations.syr,
      aes(y=reorder(asylum_country, childrenProp), x=childrenProp)) +
     stat_gradientinterval() +
  facet_wrap(~ neighbor, scales = "free")

### by region of asylum
t.dem.asyreg.children <- dem_longMissing %>% 
  filter(!is.na(children)) %>%
  group_by(asylum_region) %>% 
  summarise(children = sum(children), total = sum(total)) %>% 
  mutate(childrenProp = children/total)

t.dem.asyreg.female <- dem_longMissing %>% 
  filter(!is.na(female)) %>%
  group_by(asylum_region) %>% 
  summarise(female = sum(female), total = sum(total)) %>% 
  mutate(femaleProp = female/total)

imputations.asyreg <- imputations %>% 
  group_by(imputation, asylum_region) %>% 
  summarise(children = sum(children), female = sum(female), total = sum(total)) %>% 
  ungroup() %>%
  mutate(childrenProp = children/total,
         femaleProp = female/total) %>% 
  ungroup()

t.imp.asyreg.children <- imputations.asyreg %>% 
  group_by(asylum_region) %>% 
  summarise(childrenMeanProp = mean(childrenProp), 
            lowerUI = quantile(childrenProp, probs = 0.025), 
            upperUI = quantile(childrenProp, probs = 0.975)
            ) %>% 
  left_join(t.dem.asyreg.children %>% select(asylum_region, childrenProp), by = "asylum_region", suffix = c(".imputed", ".availableData"))

t.imp.asyreg.female <- imputations.asyreg %>% 
  group_by(asylum_region) %>% 
  summarise(femaleProp = mean(femaleProp), lowerUI = quantile(femaleProp, probs = 0.025), upperUI = quantile(femaleProp, probs = 0.975)) %>% 
  left_join(t.dem.asyreg.female %>% select(asylum_region, femaleProp), by = "asylum_region", suffix = c(".imputed", ".availableData"))

### world
t.dem.children <- dem_longMissing %>% 
  filter(!is.na(children)) %>%
  summarise(children = sum(children), total = sum(total)) %>% 
  mutate(childrenProp = children/total)

t.dem.female <- dem_longMissing %>% 
  filter(!is.na(female)) %>%
  summarise(female = sum(female), total = sum(total)) %>% 
  mutate(femaleProp = female/total)

t.imp.children <- imputations %>% 
  group_by(imputation) %>% 
  summarise(children = sum(children), female = sum(female), total = sum(total)) %>% 
  mutate(childrenProp = children/total,
         femaleProp = female/total) %>% 
  ungroup() %>%
summarise(childrenMeanProp = mean(childrenProp), 
              lowerUI = quantile(childrenProp, probs = 0.025), 
              upperUI = quantile(childrenProp, probs = 0.975)
    )
